---
title: "`learningtower`: Exploring standardised test scores across the globe"
abstract: >
  An abstract of less than 150 words - Discuss what the paper talks about with a little introduction.
draft: true
author: 
  # see ?rjournal_article for more information
  - name: Priya Ravindra Dingorkar
    url: https://www.linkedin.com/in/priya-dingorkar/
    email: priyadingorkar@gmail.com
    affiliation: Monash University
    address:
    - Department Econometrics and Business Statistics
    - Clayton, Australia
  - name: Kevin Wang
    affiliation: University of Sydney
    address:
    - School of Mathematics and Statistics 
    - Sydney, Australia
    url: https://kevinwang09.github.io/
    email: kevinwangstats@gmail.com 
  - name: Di Cook
    affiliation: Monash University
    address:
    - Department Econometrics and Business Statistics
    - Clayton, Australia
    url: http://dicook.org/
    email: dicook@monash.edu
type: package
creative_commons: CC BY
output: 
  rjtools::rjournal_web_article
bibliography: learningtower.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
library(rjtools)
library(plotly)
library(ggplot2)
library(palmerpenguins)
```

# Introduction

The Organization for Economic Cooperation and Development [OECD](OECD - https://www.oecd.org/about/) is a global organization that aims to create better policies for better lives. Their mission is to create policies that promote prosperity, equality, opportunity, and well-being for all. [PISA](PISA - https://www.oecd.org/pisa/) is one of OECD's Programme for International Student Assessment. PISA assesses 15-year-olds potential to apply their knowledge and abilities in reading, mathematics, and science to real-world challenges. OECD launched this in 1997, it was initially administered in 2000, and it currently includes over [80 nations](https://www.oecd.org/pisa/aboutpisa/pisa-participants.htm). The PISA study, conducted every three years, provides comparative statistics on 15-year-olds performance in reading, maths, and science. This paper describes how to utilize the `learningtower` package, which offers OECD PISA datasets from 2000 to 2018 in an easy-to-use format. This dataset comprises information on their test results and other socioeconomic factors, as well as information on their school and its infrastructure and the countries participating in the program.

# What is PISA?

PISA assesses the extent to which children approaching the end of compulsory school have learned some of the information and abilities required for full participation in modern society, notably in maths, reading, and science. The examination focuses on reading, mathematics, science, and problem solving. It also assesses students capacity to replicate information and extrapolate from what they have learned and apply that knowledge in unexpected circumstances, both inside and outside of school. This approach reflects the fact that individuals are rewarded in modern economies not for what they know, but for what they can accomplish with what they know. 

This evaluation which is carried out every three years, assists in identifying students development of knowledge and skills throughout the world, which can provide actionable insights and therefore assist education policymakers. PISA is well known for its distinctive testing characteristics, which include policy orientation, an innovative notion of literacy, relevance to lifelong learning, regularity, and breadth of coverage. PISA is now used as an assessment tool in many regions around the world. In addition to OECD member countries, the survey has been or is being conducted in East, South and Southeast Asia, Central, Mediterranean and Eastern Europe, and Central Asia, The Middle East, Central and South America and Africa. 

For each PISA, one domain is thoroughly examined. In 2018, for example, reading was assessed alongside mathematics and science as minor areas of assessment. The 2012 survey concentrates on mathematics, with reading, science, and problem solving serving as minor evaluation topics. PISA targets a certain age group of pupils in order to properly compare student performance worldwide. PISA students are aged between 15 years 3 months and 16 years 2 months at the time of the assessment, and have completed at least 6 years of formal schooling. They can enroll in any sort of institution, participate in full-time or part-time education, academic or vocational programs, and attend public, private, or international schools inside the country. Using this age across nations and throughout time allows PISA to compare the knowledge and abilities of people born in the same year who are still in school at the age of 15, irrespective of their diverse schooling.

The PISA test is primarily computer-based and lasts around 2 hours. The questions are a mix of multiple choice and free entry. A few countries that were not ready for computer-based delivery carried out the testing on paper. Each student may have a unique set of questions. An example of the test may be seen [here.](https://www.oecd.org/pisa/test/) PISA assessment areas seek to measure the following aspects of students' literacy in math, reading, and science. The goal of mathematical literacy is to assess students ability to grasp and interpret mathematics in a variety of settings. Reading literacy assesses students' capacity to absorb, apply, analyze, and reflect on texts in order to attain required goals and participate in society. Science literacy is described as the ability to engage with science-related issues and scientific concepts as a reflective citizen.

PISA data is publicly accessible for [download](https://www.oecd.org/pisa/data/). Furthermore, reading the [data documentation](https://www.oecd.org/pisa/data/pisa2018technicalreport/Ch.09-Scaling-PISA-Data.pdf) reveals that the disclosed PISA scores are generated using a sophisticated linear model applied to the data. For each student, several values are simulated. This is known as synthetic data, and it is a popular technique to ensuring data privacy. The data can still be deemed accurate within the mean, variance, and stratum used in the original data's model. In addition, if you continue to obtain data from the PISA website, you will be needed to download huge files in SPSS format in order to receive the results from the PISA website. This has now been simplified into a single R package called `learningtower`, which contains all of the PISA scores from the years 2000 to 2018.

# Data Compilation


# R package - `learningtower`

['learningtower'](https://cran.r-project.org/web/packages/learningtower/index.html) is an easy-to-use R package that provides quick access to a variety of variables using OECD PISA data collected over a three-year period from 2000 to 2018. This dataset includes information on on the PISA test scores in mathematics, reading, and science. Furthermore, these datasets include information on other socioeconomic aspects, as well as information on their school and its facilities, as well as the nations participating in the program. 


The motivation for developing the 'learningtower' package was sparked by the announcement of the PISA 2018 results, which caused a collective wringing of hands in the Australian press, with headlines such as ["Vital Signs: Australia's slipping student scores will lead to greater income inequality"](https://theconversation.com/vital-signs-australias-slipping-student-scores-will-lead-to-greater-income-inequality-128301) and ["In China, Nicholas studied maths 20 hours a week. In Australia, it's three"](https://www.smh.com.au/education/in-china-nicholas-studied-maths-20-hours-a-week-in-australia-it-s-three-20191203-p53ggv.html).That's when several academics from Australia, New Zealand, and Indonesia decided to make things easier by providing easy access to PISA scores as part of the [ROpenSci OzUnconf](https://ozunconf19.ropensci.org/), which was held in Sydney from December 11 to 13, 2019. The data from this survey, as well as all other surveys performed since the initial collection in 2000, is freely accessible to the public. However, downloading and loading data in R environment using SPSS formats could could be time consuming. As a result, we have made a more convenient subset of the data freely available in a new R package called 'learningtower,' along with sample code for analysis. 

The learningtower package is primarily comprised of three datasets: `student`, `school`, and `countrycode.` The `student` dataset includes results from triennial testing of 15-year-olds throughout the world. This dataset also includes information about their parents' education, family wealth, gender, and presence of computers, internet, vehicles, books, rooms, desks, and other comparable factors. Keeping the package's size restrictions in mind. The package only comprises a subset of student data, known as the `student subset data`; this subset of student data is accessible for all years in which the PISA experiment was conducted.The full year wise student data set can be downloaded using the `load_student()` function included in this [package.](https://kevinwang09.github.io/learningtower/) The school dataset includes school weight as well as other information such as school funding distribution, whether the school is private or public, enrolment of boys and girls, school size, and similar other characteristics of interest of different schools these 15-year-olds attend around the world. The countrycode dataset includes a mapping of a country's ISO code to its name.







# Example Analysis

# Gender Analysis

# Socioeconomic Factors Analysis 

# Temoral Trend Australia 

# Discussion

# Testing

Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:penguins-plotly)', '\\@ref(fig:penguins-ggplot)'))` shows an `r ifelse(knitr::is_html_output(), 'interactive', '')` plot of the \CRANpkg{palmerpenguins} data [@palmerpenguins], made using the `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\CRANpkg{plotly}', '\\CRANpkg{ggplot2}'))` package. This data features three penguin species which has a lovely illustration by Alison Horst in Figure \@ref(fig:penguins-alison).

```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name.", include=knitr::is_html_output(), eval=knitr::is_html_output()}
p <- penguins %>%
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm,
             color = species)) +
  geom_point()
ggplotly(p)
```

```{r penguins-ggplot, echo = TRUE, fig.height = 5, fig.cap="A basic non-interactive plot made with the ggplot2 package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. Visit the online article to access the interactive version made with the plotly package.", include=knitr::is_latex_output(), eval=knitr::is_latex_output()}
penguins %>%
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm,
             color = species)) +
  geom_point()
```

@RJ-2021-050

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.
