---
title: "R package learningtower: Exploring standardised test scores across the globe"
abstract: >
  An abstract of less than 150 words - Discuss what the paper talks about with a little introduction.
draft: true
author: 
  # see ?rjournal_article for more information
  - name: Priya Ravindra Dingorkar
    url: https://www.linkedin.com/in/priya-dingorkar/
    email: priyadingorkar@gmail.com
    affiliation: Monash University
    address:
    - Department Econometrics and Business Statistics
    - Clayton, Australia
  - name: Kevin Wang
    affiliation: University of Sydney
    address:
    - Data scientist and statistician
    - CSL Behring, Australia
    url: https://kevinwang09.github.io/
    email: kevinwangstats@gmail.com 
  - name: Di Cook
    affiliation: Monash University
    address:
    - Department Econometrics and Business Statistics
    - Clayton, Australia
    url: http://dicook.org/
    email: dicook@monash.edu
type: package
creative_commons: CC BY
output: 
  rjtools::rjournal_web_article
bibliography: learningtower.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(rjtools)
library(plotly)
library(ggplot2)
library(palmerpenguins)
```

# Introduction

The Organization for Economic Cooperation and Development (OECD) is a global organization that aims to create better policies for better lives. Their mission is to create policies that promote prosperity, equality, opportunity, and well-being for all.

PISA is one of the OECD's Programme for International Student Assessment. PISA assesses 15-year-olds potential to apply their knowledge and abilities in reading, mathematics, and science to real-world challenges. The OECD launched this in 1997, it was initially administered in 2000, and it currently includes over [80 nations](https://www.oecd.org/pisa/aboutpisa/pisa-participants.htm). The PISA study, conducted every three years, provides comparative statistics on 15-year-olds performance in reading, maths, and science.

## What is PISA?



## R package - learningtower

# Data Description

## Data Compilation

## Data Structure

# Exploratory Data Analysis

## Gender Analysis

## Socioeconomic Factors Analysis 

## Temoral Trend Australia 

# Future Developments

# Summary

Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:penguins-plotly)', '\\@ref(fig:penguins-ggplot)'))` shows an `r ifelse(knitr::is_html_output(), 'interactive', '')` plot of the \CRANpkg{palmerpenguins} data [@palmerpenguins], made using the `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\CRANpkg{plotly}', '\\CRANpkg{ggplot2}'))` package. This data features three penguin species which has a lovely illustration by Alison Horst in Figure \@ref(fig:penguins-alison). 

```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name.", include=knitr::is_html_output(), eval=knitr::is_html_output()}
p <- penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
ggplotly(p)
```

```{r penguins-ggplot, echo = TRUE, fig.height = 5, fig.cap="A basic non-interactive plot made with the ggplot2 package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. Visit the online article to access the interactive version made with the plotly package.", include=knitr::is_latex_output(), eval=knitr::is_latex_output()}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
```

@RJ-2021-050

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.
